Hilbert space methods to approximate Gaussian processes using Stan

### 2019-04-20 Paper/manuscript/manuscript_05.tex
improved figures

\begin{figure}
\centering
\subfigure{\includegraphics[scale=0.30, trim = 10mm 95mm 10mm 30mm, clip]{fig4_legend.pdf}}\\
\vspace{-1mm}
\subfigure{\includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig4_Post_part1.pdf}}
\subfigure{\includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig4_Cov_part1.pdf}}\\
\vspace{-3mm}
\subfigure{\includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig4_Post_part2.pdf}}
\subfigure{\includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig4_Cov_part2.pdf}}\\
\vspace{-3mm}
\subfigure{\includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig4_Post_part3.pdf}}
\subfigure{\includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig4_Cov_part3.pdf}}\\
\vspace{-3mm}
\subfigure{\includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig4_Post_part4.pdf}}
\subfigure{\includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig4_Cov_part4.pdf}}\\
\vspace{-3mm}
\subfigure{\includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig4_Post_part5.pdf}}
\subfigure{\includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig4_Cov_part5.pdf}}\\
\vspace{-3mm}
\subfigure{\includegraphics[scale=0.25, trim = 0mm 10mm 10mm 20mm, clip]{fig4_Post_part6.pdf}}
\subfigure{\includegraphics[scale=0.25, trim = 0mm 10mm 10mm 20mm, clip]{fig4_Cov_part6.pdf}}\\
\vspace{-3mm}
\caption{(left) Posterior predictive mean functions of both regular GP model and HSGP model for different boundary factors $c$. (right) Covariance functions of both regular GP model and HSGP model for different boundary factors $c$.}
  \label{fig4_Post_part1}
\end{figure}

\begin{figure}
\centering
\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig5_MSE_vs_J.pdf}}
\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig5_MSE_vs_c.pdf}}
\caption{Root mean square error (RMSE) of the proposed HSGP models computed against the regular GP model. (left) RMSE versus the number of basis functions $m$ and for different values of the boundary factor $c$. (right) RMSE versus the boundary factor $c$ and for different values of the number of basis functions $m$. }
  \label{fig5_MSE_vs_J}
\end{figure}

\begin{figure}
\centering
\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig6_lscale_vs_J.pdf}}
\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig6_magnitud_vs_J.pdf}}
\subfigure{\includegraphics[scale=0.40, trim = 22mm 45mm 120mm 0mm, clip]{fig6_legend.pdf}}
\caption{Estimated lengthscale (left) and marginal variance (right) parameters of both regular GP and HSGP models, plotted versus the number of basis functions $m$ and for different values of the boundary factor $c$.}
  \label{fig6_lscale_vs_J}
\end{figure}

@@

\begin{figure}
\centering
\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig3_lscale_vs_J_vs_c.pdf}}
\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig3_lscale_vs_J_vs_c_zoomin.pdf}}
\caption{Relation among the minimum number of basis functions $m$, the boundary factor $c$ ($c = \frac{L}{S}$), and the lengthscale normalized by the half-range of the data ($\frac{\ell}{S}$). The right-side plot is a zoom in of the left-side plot.}
  \label{fig3_lscale_vs_J_vs_c}
\end{figure}

@@

\begin{table}
\begin{flushleft}
\begin{tabular}{ c c c c}
\includegraphics[scale=0.28, trim = 0mm 30mm 10mm 20mm, clip]{fig8_Tlscale_vs_Elscale_part1.pdf} \vspace{0mm} & \hspace{-5mm} \includegraphics[scale=0.28, trim = 22mm 30mm 10mm 20mm, clip]{fig8_Tlscale_vs_Elscale_part2.pdf} \vspace{0mm} & \hspace{-5mm} \includegraphics[scale=0.28, trim = 22mm 30mm 10mm 20mm, clip]{fig8_Tlscale_vs_Elscale_part3.pdf} \vspace{0mm} & \hspace{-8mm} \multirow{3}{1em}{\vspace{-9mm} \includegraphics[scale=0.35, trim = 10mm 105mm 130mm 50mm, clip]{fig8_legend.pdf}}\\ 
\includegraphics[scale=0.28, trim = 0mm 10mm 10mm 20mm, clip]{fig8_Tlscale_vs_Elscale_part4.pdf} \vspace{0mm}  & \hspace{-5mm} \includegraphics[scale=0.28, trim = 22mm 10mm 10mm 20mm, clip]{fig8_Tlscale_vs_Elscale_part5.pdf} \vspace{0mm} & \hspace{-5mm} \includegraphics[scale=0.28, trim = 22mm 10mm 10mm 20mm, clip]{fig8_Tlscale_vs_Elscale_part6.pdf} \vspace{0mm}
\end{tabular}
\end{flushleft}
\caption{The X-axes represent the true data-generative functional lengthscales of every dataset scenario. The Y-axes represent the lengthscales estimates of GP and HSGP models. The use of different $m$ in the HSGP models is represented in separated plots. The dashed black line represents the recommended minimum lengthscales that can be closely approximate by the HSGP model in every case according to Figure \ref{fig3_lscale_vs_J_vs_c}.}
  \label{fig8_Tlscale_vs_Elscale}
\end{table}

\begin{figure}
\centering
\includegraphics[scale=0.45]{fig9_MSE_varing_lscale.pdf}
\caption{RMSE of the HSGP models for the different datasets with different wiggly effects (lengthscales) and number of basis functions $m$.}
  \label{fig9_MSE_varing_lscale}
\end{figure}

@@

\subsection{Simulated data}

@@

\begin{figure}[H]
\centering
\includegraphics[scale=0.35]{fig10_Posteriors_exI.pdf}
\caption{Posterior distributions of the proposed HSGP model, the regular GP model, and the Splines model.}
  \label{fig10_Posteriors_exI}
\end{figure}

@@

\begin{figure}[H]
\centering
\subfigure{\includegraphics[scale=0.350]{fig11_MSE_exI_inter.pdf}}
\subfigure{\includegraphics[scale=0.350]{fig11_MSE_exI_extra.pdf}}
\caption{Standarized mean square error (SMSE) of the different methods againts the true function. (left) SMSE for interpolation. (right) SMSE for extrapolation}
  \label{fig11_MSE_exI_inter}
\end{figure}

\subsection{Gay data}
This data set relates the proportion of support for same-sex marriage to the age. The data consists of 74 observations of the amount of people $y_i$ supporting same-sex marriage from a population $n_i$ per age group $i$ ($i=1,\cdots,74$). The observational model is a binomial model with parameters population $n_i$ and probability of supporting same-sex marriage $p_i$ per age group $i$,
%
\begin{equation*}
y_i \sim \text{Binomial}(p_i, n_i).
\end{equation*}

@@

\begin{figure}[H]
\centering
\subfigure{\includegraphics[scale=0.40]{fig13_MSE_train_BF_gaydata.pdf}}
\subfigure{\includegraphics[scale=0.40]{fig13_MSE_pred_BF_gaydata.pdf}}
\caption{Mean square error (MSE) of the different methods againts the true function. (left) MSE for training data. (right) MSE for test data.}
  \label{fig13_MSE_train_BF_gaydata}
\end{figure}

@@

\begin{figure}[H]
\centering
\subfigure{\includegraphics[scale=0.40]{fig14_MSE_train_gaydata.pdf}}
\subfigure{\includegraphics[scale=0.40]{fig14_MSE_pred_gaydata.pdf}}
\caption{Mean square error (MSE) of the different methods againts the true function. (left) MSE for training data. (right) MSE for test data.}
  \label{fig14_MSE_train_gaydata}
\end{figure}

\subsection{Birthday data}

\section{Multivariate examples}
\subsection{Simulated data}

@@

\begin{figure}
\centering
\subfigure{\includegraphics[scale=0.40, trim = 0mm 33mm 35mm 20mm, clip]{fig15_truefun_exII.pdf}}
\subfigure{\includegraphics[scale=0.40, trim = 20mm 33mm 35mm 20mm, clip]{fig15_gpfun_exII.pdf}}
\subfigure{\includegraphics[scale=0.40, trim = 0mm 22mm 35mm 20mm, clip]{fig15_bffun_exII.pdf}}
\subfigure{\includegraphics[scale=0.40, trim = 20mm 22mm 35mm 20mm, clip]{fig15_spfun_exII.pdf}}
\subfigure{\includegraphics[scale=0.5, trim = 0mm 5mm 0mm 155mm, clip]{fig15_colorbar_exII.pdf}}
\caption{(a) True data-generative function. (b) Posterior mean function of the GP model. (c) Posterior mean function of the HSGP model. (d) Posterior mean function of the splines (SP) model.}
  \label{fig15_Posteriors_exII}
\end{figure}

\begin{figure}
\centering
\subfigure{\includegraphics[scale=0.36, trim = 8mm 23mm 40mm 20mm, clip]{fig16_errorGP_exII.pdf}}
\subfigure{\includegraphics[scale=0.36, trim = 20mm 23mm 40mm 20mm, clip]{fig16_errorHSGP_exII.pdf}}
\subfigure{\includegraphics[scale=0.36, trim = 20mm 23mm 40mm 20mm, clip]{fig16_errorSP_exII.pdf}}
\subfigure{\includegraphics[scale=0.34, trim = 150mm 20mm 10mm 20mm, clip]{fig16_error_colorbar_exII.pdf}}
\caption{Mean error between the data-generative function and the GP (a), HSGP (b) and splines (c) models.}
  \label{fig16_errors_exII}
\end{figure}

@@

\subsection{Case III: Diabetes data}\label{sec:bf_caseIV}
\subsection{Case IV: Leukemia data}\label{sec:bf_caseV}
\subsection{Case V: Land use spatio-temporal classification task}\label{sec:bf_caseVII}

>>>>

\begin{figure}
\begin{center}
\begin{tabular}{ c c c}
\includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig3_Post_part1.pdf} & \hspace{-5mm} \includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig3_Cov_part1.pdf} & \hspace{-6mm} \multirow{30}{1em}{ \includegraphics[scale=0.35, trim = 20mm 60mm 120mm 40mm, clip]{fig3_legend.pdf}}\\ 
\includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig3_Post_part2.pdf} & \hspace{-5mm} \includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig3_Cov_part2.pdf} &\\
\includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig3_Post_part3.pdf} & \hspace{-5mm} \includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig3_Cov_part3.pdf} &\\
\includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig3_Post_part4.pdf} & \hspace{-5mm} \includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig3_Cov_part4.pdf} & \\
\includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig3_Post_part5.pdf} & \hspace{-5mm} \includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig3_Cov_part5.pdf} &\\
\includegraphics[scale=0.25, trim = 0mm 5mm 10mm 20mm, clip]{fig3_Post_part6.pdf} & \hspace{-5mm} \includegraphics[scale=0.25, trim = 0mm 5mm 10mm 20mm, clip]{fig3_Cov_part6.pdf} &
\end{tabular}
\end{center}
\caption{The X-axes represent the true data-generative functional lengthscales of every dataset scenario. The Y-axes represent the lengthscales estimates of GP and HSGP models. The use of different $m$ in the HSGP models is represented in separated plots. The dashed black line represents the recommended minimum lengthscales that can be closely approximate by the HSGP model in every case according to Figure \ref{fig3_lscale_vs_J_vs_c}.}
  \label{fig3_Post_part1}
\end{figure}

\begin{figure}
\centering
\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig4_MSE_vs_J.pdf}}
\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig4_MSE_vs_c.pdf}}
\caption{Root mean square error (RMSE) of the proposed HSGP models computed against the regular GP model. (left) RMSE versus the number of basis functions $m$ and for different values of the boundary factor $c$. (right) RMSE versus the boundary factor $c$ and for different values of the number of basis functions $m$. }
  \label{fig4_MSE_vs_J}
\end{figure}

\begin{figure}
\centering
\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig5_lscale_vs_J.pdf}}
\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig5_magnitud_vs_J.pdf}}
\subfigure{\includegraphics[scale=0.40, trim = 22mm 45mm 120mm 0mm, clip]{fig5_legend.pdf}}
\caption{Estimated lengthscale (left) and marginal variance (right) parameters of both regular GP and HSGP models, plotted versus the number of basis functions $m$ and for different values of the boundary factor $c$.}
  \label{fig5_lscale_vs_J}
\end{figure}

@@

\begin{figure}
\centering
\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig6_lscale_vs_J_vs_c.pdf}}
\hspace{3mm}
\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig6_lscale_vs_J_vs_c_zoomin.pdf}}
\caption{Relation among the minimum number of basis functions $m$, the boundary factor $c$ ($c = \frac{L}{S}$), and the lengthscale normalized by the half-range of the data ($\frac{\ell}{S}$). The right-side plot is a zoom in of the left-side plot.}
  \label{fig6_lscale_vs_J_vs_c}
\end{figure}

@@

\begin{figure}
\begin{flushleft}
\begin{tabular}{ c c c c}
\includegraphics[scale=0.28, trim = 0mm 30mm 10mm 20mm, clip]{fig8_Tlscale_vs_Elscale_part1.pdf} \vspace{0mm} & \hspace{-5mm} \includegraphics[scale=0.28, trim = 22mm 30mm 10mm 20mm, clip]{fig8_Tlscale_vs_Elscale_part2.pdf} \vspace{0mm} & \hspace{-5mm} \includegraphics[scale=0.28, trim = 22mm 30mm 10mm 20mm, clip]{fig8_Tlscale_vs_Elscale_part3.pdf} \vspace{0mm} & \hspace{-8mm} \multirow{3}{1em}{\vspace{-9mm} \includegraphics[scale=0.35, trim = 10mm 105mm 130mm 45mm, clip]{fig8_legend.pdf}}\\ 
\includegraphics[scale=0.28, trim = 0mm 10mm 10mm 20mm, clip]{fig8_Tlscale_vs_Elscale_part4.pdf} \vspace{0mm}  & \hspace{-5mm} \includegraphics[scale=0.28, trim = 22mm 10mm 10mm 20mm, clip]{fig8_Tlscale_vs_Elscale_part5.pdf} \vspace{0mm} & \hspace{-5mm} \includegraphics[scale=0.28, trim = 22mm 10mm 10mm 20mm, clip]{fig8_Tlscale_vs_Elscale_part6.pdf} \vspace{0mm}
\end{tabular}
\end{flushleft}
\caption{The X-axes represent the true data-generative functional lengthscales of every dataset scenario. The Y-axes represent the lengthscales estimates of GP and HSGP models. The use of different $m$ in the HSGP models is represented in separated plots. The dashed black line represents the recommended minimum lengthscales that can be closely approximate by the HSGP model in every case according to Figure \ref{fig6_lscale_vs_J_vs_c}.}
  \label{fig8_Tlscale_vs_Elscale}
\end{figure}

\begin{figure}
\centering
\includegraphics[scale=0.35]{fig9_MSE_varing_lscale.pdf}
\caption{RMSE of the HSGP models for the different datasets with different wiggly effects (lengthscales) and number of basis functions $m$.}
  \label{fig9_MSE_varing_lscale}
\end{figure}

@@

\subsection{Study case I: Simulated data}

@@

\begin{figure}
\centering
\includegraphics[width=\textwidth]{fig10_Posteriors_exI.pdf}
\caption{Posterior distributions of the proposed HSGP model, the regular GP model, and the Splines model.}
  \label{fig10_Posteriors_exI}
\end{figure}

@@

\begin{figure}
\centering
\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig11_MSE_exI_inter.pdf}}
\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig11_MSE_exI_extra.pdf}}
\subfigure{\includegraphics[scale=0.40, trim = 22mm 60mm 120mm 0mm, clip]{fig11_MSE_exI_legend.pdf}}
\caption{Standarized root mean square error of the different methods againts the true function. (left) SMSE for interpolation. (right) SMSE for extrapolation.}
  \label{fig11_MSE_exI_inter}
\end{figure}

\subsection{Study case II: Gay data}
This data set relates the proportion of support for same-sex marriage to the age. The data consists of 74 observations of the amount of people $y_i$ supporting same-sex marriage from a population $n_i$ per age group $i$ ($i=1,\cdots,74$). The observational model is a binomial model with parameters population $n_i$ and probability of supporting same-sex marriage $p_i$ per age group $i$,
%
\begin{equation*}
y_i \sim \text{Binomial}(p_i, n_i).
\end{equation*}

@@

\begin{figure}
\centering
\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig13_MSE_train_BF_gaydata.pdf}}
\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig13_MSE_pred_BF_gaydata.pdf}}
\subfigure{\includegraphics[scale=0.40, trim = 22mm 55mm 120mm 0mm, clip]{fig13_legend_gaydata.pdf}}
\caption{Root mean square error (RMSE) of the different methods against the true function. (left) RMSE for training data. (right) RMSE for test data.}
  \label{fig13_MSE_train_BF_gaydata}
\end{figure}

@@

\begin{figure}
\centering
\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig14_MSE_train_gaydata.pdf}}
\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig14_MSE_pred_gaydata.pdf}}
\subfigure{\includegraphics[scale=0.40, trim = 22mm 50mm 120mm 0mm, clip]{fig14_legend_gaydata.pdf}}
\caption{Mean square error (MSE) of the different methods againts the actual data. (left) MSE for training data. (right) MSE for test data.}
  \label{fig14_MSE_train_gaydata}
\end{figure}

%\begin{figure}[H]
%\centering
%\subfigure{\includegraphics[scale=0.40]{fig14_MSE_train_gaydata.pdf}}
%\subfigure{\includegraphics[scale=0.40]{fig14_MSE_pred_gaydata.pdf}}
%\caption{Mean square error (MSE) of the different methods againts the actual data. (left) MSE for training data. (right) MSE for test data.}
%  \label{fig14_MSE_train_gaydata}
%\end{figure}

\subsection{Study case III: Birthday data}

\vspace{5mm}
\section{Multivariate examples}
\subsection{Study case IV: Simulated data}

@@

\begin{figure}
\begin{center}
\begin{tabular}{ c c c}
\includegraphics[scale=0.35, trim = 0mm 33mm 35mm 20mm, clip]{fig15_truefun_exII.pdf} & \hspace{-5mm} \includegraphics[scale=0.35, trim = 20mm 33mm 35mm 20mm, clip]{fig15_gpfun_exII.pdf} & \hspace{-5mm}\multirow{-5.5}{*}{ \includegraphics[scale=0.35, trim = 150mm 20mm 10mm 10mm, clip]{fig15_colorbar_vertical_exII.pdf}}\\ 
\includegraphics[scale=0.35, trim = 0mm 22mm 35mm 20mm, clip]{fig15_bffun_exII.pdf}  & \hspace{-5mm} \includegraphics[scale=0.35, trim = 20mm 22mm 35mm 20mm, clip]{fig15_spfun_exII.pdf}
\end{tabular}
\end{center}
\caption{(a) True data-generative function. (b) Posterior mean function of the GP model. (c) Posterior mean function of the HSGP model. (d) Posterior mean function of the splines (SP) model.}
  \label{fig15_Posteriors_exII}
\end{figure}

%\begin{figure}
%\centering
%\subfigure{\includegraphics[scale=0.40, trim = 0mm 33mm 35mm 20mm, clip]{fig15_truefun_exII.pdf}}
%\subfigure{\includegraphics[scale=0.40, trim = 20mm 33mm 35mm 20mm, clip]{fig15_gpfun_exII.pdf}}
%\subfigure{\includegraphics[scale=0.40, trim = 0mm 22mm 35mm 20mm, clip]{fig15_bffun_exII.pdf}}
%\subfigure{\includegraphics[scale=0.40, trim = 20mm 22mm 35mm 20mm, clip]{fig15_spfun_exII.pdf}}
%\subfigure{\includegraphics[scale=0.5, trim = 0mm 5mm 0mm 155mm, clip]{fig15_colorbar_exII.pdf}}
%\caption{(a) True data-generative function. (b) Posterior mean function of the GP model. (c) Posterior mean function of the HSGP model. (d) Posterior mean function of the splines (SP) model.}
%  \label{fig15_Posteriors_exII}
%\end{figure}

\begin{figure}
\centering
\subfigure{\includegraphics[scale=0.35, trim = 8mm 23mm 40mm 20mm, clip]{fig16_errorGP_exII.pdf}}
\subfigure{\includegraphics[scale=0.35, trim = 20mm 23mm 40mm 20mm, clip]{fig16_errorHSGP_exII.pdf}}
\subfigure{\includegraphics[scale=0.35, trim = 20mm 23mm 40mm 20mm, clip]{fig16_errorSP_exII.pdf}}
\subfigure{\includegraphics[scale=0.33, trim = 149mm 20mm 5mm 20mm, clip]{fig16_error_colorbar_exII.pdf}}
\caption{Mean error between the data-generative function and the GP (a), HSGP (b) and splines (c) models.}
  \label{fig16_errors_exII}
\end{figure}

@@

\subsection{Study case V: Diabetes data}\label{sec:bf_caseV}
The next example presents an epidemiological study of the diabetes diseases. The study aims to relate the probability of suffering from diabetes to some risk factors. The data set consists of 392 individuals ($i=1,\cdots,n=392$) from which the binary variable of suffering ($y_i=1$) or not suffering ($y_i=0$) from diabetes have been observed. The matrix $X=\{\bm{x}_i\}_{i=1}^{n=392} \in {\rm I\!R}^{n\times 3}$,  with $\bm{x}_i=\{x_{i1},x_{i2},x_{i3}\} \in {\rm I\!R}^{3}$, contains the risk factor values, Glucose ($x_{i1}$), Pregnancy ($x_{i2}$) and Age ($x_{i3}$), per individual $i$,. The observational model is a bernoulli model with parameter the probability $p_i$ of suffering from diabetes per observation $i$,
%
\begin{equation*}
y_i \sim \text{Bernoulli}(p_i).
\end{equation*}

\noindent The goal is to estimate the 
 probability $p_i$ as a function of the risk factors, which function $f(X)$ is modeled as a Gaussian process with a squared exponential covariance function $k$ depending of the matrix $X$ of risk factors and hyperparameters $\theta=\{\alpha,\ell\}$, and related to the probabilities $p_i$ through the {\it logit} link function,
%
\begin{eqnarray*} \label{eq:gpprior_gay}
p_i &=& \text{logit}(f(\bm{x}_i)) \nonumber \\
f(X) &\sim& \mathcal{GP}(0, k(X,X, \theta).
\end{eqnarray*}

\noindent The hyperparameters $\alpha$ and $\ell$ represent the marginal variance and lengthscale of the GP process. Notice that a scalar lengthscale is considered only in the multivariate covariance function.

In the HSGP model with $D$ input dimensions, the latent function $f$, evaluated at input vector $\bm{x} \in {\rm I\!R}^D$, is approximated as (\ref{approxf_multi}),
%
\begin{equation*}
f(\bm{x}) \approx \sum_{j}^m \left( S(\sqrt{\bm{\lambda}_j})\right)^{1/2} \phi_j(\bm{x}) \beta_j,
\end{equation*}

\noindent where $\bm{\lambda}_j$ is the $D$-vector (\ref{eigenvalue_multi}) with elements that are univariate eigenvalues (\ref{eigenvalue}) whose indices correspond to the elements of the $D$-tupple $\mathbb{S}_{j\cdotp}$,
%
\begin{align*}
\bm{\lambda}_j &= \left\{ \lambda_{\mathbb{S}_{jd}} \right\}_{d=1}^{D} = \left\{ \left(\tfrac{\pi \mathbb{S}_{jd}}{2L_d}\right)^2 \right\}_{d=1}^{D},
\end{align*}

\noindent and $\phi_j$ is the multivariate eigenfunction (\ref{eigenfunction_multi}), as the product of univariate eigenfunctions (\ref{eigenfunction}) whose indices correspond to the elements of the $D$-tupple $\mathbb{S}_{j\cdotp}$,
%
\begin{align*}
\phi_j(\bm{x}) &= \prod_{d=1}^{D} \phi_{\mathbb{S}_{jd}} = \prod_{d=1}^{D} \sqrt{\frac{1}{L_d}} \text{sin}\left(\sqrt{\lambda_{\mathbb{S}_{jd}}}(x_d+L_d)\right),
\end{align*}

\noindent where $x_d$ is the input value corresponding to dimension $d$. $S$ is the spectral density (\ref{eq:specdens_inf}), as a function of $\sqrt{\bm{\lambda}_j}= \left\{ \sqrt{\lambda_{\mathbb{S}_{jd}}} \right\}_{d=1}^{D}$, of the $D$-dimensional square exponential covariance function with a scalar lengthscale $\ell$,
%
\begin{eqnarray*}
S(\sqrt{\bm{\lambda}_j})= \sigma^2 2\pi \ell^D  \mathrm{exp}\bigg(-\frac{1}{2} \ell^2 \sum_{d=1}^D  \sqrt{\lambda_{\mathbb{S}_{jd}}}^2 \bigg). 
\end{eqnarray*}

\noindent In the previous equations, $j$ denotes the index for the $m= \prod_{d=1}^{D} m_d$ multivariate basis functions (\ref{m_multi}), where $m_d$ is the number of basis functions considered for dimension $d$. $\mathbb{S}$ is the matrix of $D$-tuples, with rows being the indices of every possible combinations of univariate eigenvalues (\ref{eigenvalue}) over the $D$ dimensions. $L_d$ is the boundary for the dimension $d$. The parameters $\beta_j$ are $\mathcal{N}(0,1)$ distributed, and $\alpha$ and $\ell$ are the marginal variance and lengthscale, respectively, of the approximate multivariate covariance function.

In order to do model comparison, in addition to the regular GP and HSGP models, a $D$-dimensional splines-based model is also fitted using a cubic spline basis, penalized by the conventional intergrated square second derivative cubic spline penalty \citep{wood2017generalized}, and implemented in the R-package \textit{mgcv}. A Bayesian approach is used to fit this spline model using the R-package \textit{brms}.

In order to analyze the increase of computational time  of the HSGP model with higher dimensions $D$, we perform the study twice, first using two risk factors only, Glucose and Pregnancy, and second using the three risk factors, Glucose, Pregnancy and Age. With the HSGP approach in a multivariate input space, choosing the optimal number of basis functions per dimension can be crucial to avoid excessive computational times.

%Figure \ref{fig15_Posteriors_exII} shows the data-generative GP function, from where the dataset was drawn and the mean posterior functions of the three models fitted over the dataset: the regular GP, the HSGP, and the splines models. The noisy observations are also plotted in the plots as circles. For the HSGP model, $m_1=40$ and $m_2=40$ basis functions for each dimension respectively, were used, which lead to a total of 1600 multivariate basis functions. A boundary factor for each dimension $c_1=1.5$ and $c_2=1.5$ are used. For the splines model, 40 knots in each dimension were used.

\subsection{Study case VI: Leukemia data}\label{sec:bf_caseVI}
\subsection{Study case VII: Land use spatio-temporal classification task}\label{sec:bf_caseVII}
